Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows

### 2024-10-19 manuscript/manuscript.Rmd
update ms.

Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial number 1000 for the selected model.

@@

```{r plot-prime-effects, fig.cap = "(ref:descr-fig5-caption)", out.width='80%'}
knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000.png")
```

>>>>

Figure 5 shows the effects of congruent and incongruent primes relative to neutral primes, for each time bin in trial numbers 500, 1000, and 1500 for the selected model.

@@

```{r plot-prime-effects, fig.cap = "(ref:descr-fig5-caption)", out.width='80%'}
knitr::include_graphics("../Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials.png")
```

Tabel X shows the mean point estimates and the upper and lower limits of the 95% highest density intervals for each bin

```{r}
int_tab <- read_csv("../Tutorial_2_Bayesian/tables/effects_intervals_table.csv")
```

(ref:int-table-caption) Point and interval estimates.

(ref:int-table-note-caption) The posterior distributions of the effects of congruent and incongruent primes relative to no primes are summarized by point estimates and the lower and upper ...

```{r int-table}
apa_table(
  int_tab,
  caption = "(ref:int-table-caption)",
  note = "(ref:int-table-note-caption)",
  placement = "H",
  font_size = "scriptsize",
  longtable = TRUE,
  repeat_header = TRUE
)
```

### 2024-10-19 manuscript/manuscript.tex
update ms.

Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X). The equations that define both of these functions are reported in supplementary materials (part A). The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins. In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occured in the past, i.e., in one of the prior bins. This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before. In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t. In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.

@@

In the context of experimental psychology, it is common for participants to be presented with a task that has a right and a wrong answer. For example, a task may involve choosing between two response options with only one of them being correct. For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ. X in Supps), which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).

@@

We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses. Web links are printed in bold. The content of the tutorials is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).

@@

Our data wrangling procedures serve two related purposes. First, we want to summarise and visualise descriptive statistics that relate to our main research questions. Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches. The two types of data structure we label as `person-trial' data (Table 1) and `person-trial-bin' data (Table 2). The `person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment. In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial. Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred. It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occured after the timewindow of interest. Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.

@@

The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored.
One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant.

Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) - compare to Figure 1. A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), ca(t) and their estimated standard errors (se). At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.

@@

@@

To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type. As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8. Note the negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.

@@

In this second tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial. Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).

@@

In general, there are three assumptions one can make or relax when adding experimental predictor variables: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).

In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1. We select the analysis range (200,600{]} and the cloglog link. The data is prepared as follows.

@@

{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_con_incon_trial1000}

@@

@@

\noindent F(t) = 1-S(t) = P(RT \(\leq\) t). \hfill  (4)

The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins. Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t.

@@

which gives you the probability that a response is correct given that it has been emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977). This latter function is also known as the micro-level speed-accuracy tradeoff function.

When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition. Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t). The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.

@@

\begin{itemize}
\tightlist
\item
  censor(df,timeout,bin\_width)
\item
  ptb(df)
\item
  setup\_lt(ptb)
\item
  setup\_lt\_2IV(ptb)
\item
  calc\_ca(df)
\item
  calc\_ca\_2IV(df)
\item
  join\_lt\_ca(df1,df2)
\item
  join\_lt\_ca\_2IV(df1,df2)
\item
  extract\_median(df)
\item
  extract\_median\_2IV(df)
\item
  plot\_eha(df,subj,haz\_yaxis)
\item
  plot\_eha\_2IV(df,subj,haz\_yaxis)
\item
  plot\_eha\_agg(df,subj,haz\_yaxis)
\end{itemize}

@@

cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}.

@@

cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}.

>>>>

Discrete-time EHA focuses on the discrete-time hazard function and the discrete-time survivor function (Figure X). The equations that define both of these functions are reported in supplementary materials (part A). The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins. In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins. This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before. In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t. In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.

@@

In the context of experimental psychology, it is common for participants to be presented with a task that has a right and a wrong answer. For example, a task may involve choosing between two response options with only one of them being correct. For such two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function (see equ. X in Supps), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977).

@@

We used R (Version 4.4.0; R Core Team, 2024)\footnote{We, furthermore, used the R-packages \emph{bayesplot} (Version 1.11.1; Gabry, Simpson, Vehtari, Betancourt, \& Gelman, 2019), \emph{brms} (Version 2.21.0; Bürkner, 2017, 2018, 2021), \emph{citr} (Version 0.3.2; Aust, 2019), \emph{cmdstanr} (Version 0.8.1.9000; Gabry, Češnovar, Johnson, \& Bronder, 2024), \emph{dplyr} (Version 1.1.4; Wickham, François, Henry, Müller, \& Vaughan, 2023), \emph{forcats} (Version 1.0.0; Wickham, 2023a), \emph{ggplot2} (Version 3.5.1; Wickham, 2016), \emph{lme4} (Version 1.1.35.5; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lubridate} (Version 1.9.3; Grolemund \& Wickham, 2011), \emph{Matrix} (Version 1.7.0; Bates, Maechler, \& Jagan, 2024), \emph{nlme} (Version 3.1.166; Pinheiro \& Bates, 2000), \emph{papaja} (Version 0.1.2.9000; Aust \& Barth, 2023), \emph{patchwork} (Version 1.2.0; Pedersen, 2024), \emph{purrr} (Version 1.0.2; Wickham \& Henry, 2023), \emph{RColorBrewer} (Version 1.1.3; Neuwirth, 2022), \emph{Rcpp} (Eddelbuettel \& Balamuta, 2018; Version 1.0.12; Eddelbuettel \& François, 2011), \emph{readr} (Version 2.1.5; Wickham, Hester, \& Bryan, 2024), \emph{RJ-2021-048} (Bengtsson, 2021), \emph{standist} (Version 0.0.0.9000; Girard, 2024), \emph{stringr} (Version 1.5.1; Wickham, 2023b), \emph{tibble} (Version 3.2.1; Müller \& Wickham, 2023), \emph{tidybayes} (Version 3.0.6; Kay, 2023), \emph{tidyr} (Version 1.3.1; Wickham, Vaughan, \& Girlich, 2024), \emph{tidyverse} (Version 2.0.0; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.4; Barth, 2023).} for all reported analyses. The content of the tutorials is mainly based on Allison (2010), Singer and Willett (2003), McElreath (2018), Kurz (2023a), and Kurz (2023b).

@@

Our data wrangling procedures serve two related purposes. First, we want to summarise and visualise descriptive statistics that relate to our main research questions. Second, we want to produce two different datasets that can each be submitted to different types of inferential modelling approaches. The two types of data structure we label as `person-trial' data (Table 1) and `person-trial-bin' data (Table 2). The `person-trial' data will be familiar to most researchers who record behavioural responses from participants, as it represents the measured RT and accuracy per trial within an experiment. In contrast, the `person-trial-bin' data has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial. Therefore, the `person-trial-bin' dataset generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred. It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the timewindow of interest. Likewise, when the event occurs in bin 1 there would only be one row of data for that trial.

@@

The warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored.
One can now inspect different aspects, including the life table for a particular condition of a particular subject, and a plot of the different functions for a particular participant. It is important to visually inspect the functions first for each participant, in order to identify possible cheaters (e.g., a flat conditional accuracy function at .5 indicates (s)he was only guessing), outlying individulas, and/or different groups with qualitatively different behavior.

Table 3 shows the life table for condition ``blank'' (no prime stimulus presented) for participant 6 - compare to Figure 1. A life table includes for each time bin, the risk set (number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of h(t), S(t), possibly ca(t), and their estimated standard errors (se). At time point zero, no events can occur and therefore h(t) and ca(t) are undefined.

@@

When participants show qualitatively the same distributional patterns, on might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).

@@

To this end, Tutorial 1b illustrates how to calculate and plot the descriptive statistics for the full data set of Experiment 1 of Panis and Schmidt (2016), which includes two independent variables: mask type and prime type. As we use the same functional programming approach as in Tutorial 1a, we simply present the sample-based functions for participant 6 in Figure 8.

Compared to the no-mask condition (column 1 in Figure 8), there is a negative compatibility effect in the hazard and conditional accuracy functions when a (relevant, irrelevant, or lines) mask is present.
T\ldots{}

@@

In this third tutorial, we illustrate how to fit Bayesian hazard regression models to the masked response priming data set used in the first tutorial. Fitting (Bayesian or non-Bayesian) regression models to the data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).

@@

In general, there are three assumptions one can make or relax when adding experimental predictor variables and other covariates: The linearity assumption for continuous predictors (the effect of a 1 unit change is the same anywhere on the scale), the additivity assumption (predictors do not interact), and the proportionality assumption (predictors do not interact with TIME).

In this tutorial we will fit four Bayesian multilevel models (i.e., generalized linear mixed models) to the person-trial-bin oriented data set that we created in Tutorial 1a. We select the analysis range (200,600{]} and the cloglog link. The data is prepared as follows.

@@

{\centering \includegraphics[width=0.8\linewidth,height=0.67\textheight,]{../Tutorial_2_Bayesian/figures/M4effects_con_incon_3trials}

@@

Tabel X shows the mean point estimates and the upper and lower limits of the 95\% highest density intervals for each bin

\begin{center}
\begin{ThreePartTable}

\begin{TableNotes}[para]
\normalsize{\textit{Note.} The posterior distributions of the effects of congruent and incongruent primes relative to no primes are summarized by point estimates and the lower and upper \ldots{}}
\end{TableNotes}

\scriptsize{

\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:int-table}Point and interval estimates.}\\
\toprule
bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:int-table} continued}}\\
\toprule
bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
\midrule
\endhead
240.00 & c500 & 2.18 & 0.33 & 4.05 & 0.95 & 8.82\\
280.00 & c500 & 1.11 & -0.02 & 2.11 & 0.95 & 3.03\\
320.00 & c500 & 0.37 & -0.34 & 1.04 & 0.95 & 1.45\\
360.00 & c500 & -0.09 & -0.70 & 0.48 & 0.95 & 0.91\\
400.00 & c500 & -0.35 & -1.02 & 0.34 & 0.95 & 0.71\\
440.00 & c500 & -0.45 & -1.50 & 0.56 & 0.95 & 0.64\\
480.00 & c500 & -0.48 & -2.32 & 1.27 & 0.95 & 0.62\\
520.00 & c500 & -0.48 & -3.57 & 2.52 & 0.95 & 0.62\\
560.00 & c500 & -0.52 & -5.69 & 4.27 & 0.95 & 0.60\\
600.00 & c500 & -0.66 & -8.56 & 6.99 & 0.95 & 0.52\\
240.00 & c1000 & 3.03 & 1.37 & 4.82 & 0.95 & 20.63\\
280.00 & c1000 & 1.63 & 0.68 & 2.63 & 0.95 & 5.13\\
320.00 & c1000 & 0.64 & -0.02 & 1.24 & 0.95 & 1.90\\
360.00 & c1000 & -0.01 & -0.57 & 0.52 & 0.95 & 0.99\\
400.00 & c1000 & -0.38 & -1.01 & 0.22 & 0.95 & 0.68\\
440.00 & c1000 & -0.54 & -1.52 & 0.32 & 0.95 & 0.58\\
480.00 & c1000 & -0.54 & -2.20 & 1.11 & 0.95 & 0.58\\
520.00 & c1000 & -0.45 & -3.40 & 2.35 & 0.95 & 0.64\\
560.00 & c1000 & -0.34 & -5.78 & 3.90 & 0.95 & 0.71\\
600.00 & c1000 & -0.25 & -8.34 & 6.73 & 0.95 & 0.78\\
240.00 & c1500 & 3.05 & 1.07 & 5.02 & 0.95 & 21.02\\
280.00 & c1500 & 1.54 & 0.40 & 2.65 & 0.95 & 4.66\\
320.00 & c1500 & 0.42 & -0.36 & 1.13 & 0.95 & 1.52\\
360.00 & c1500 & -0.38 & -1.05 & 0.21 & 0.95 & 0.68\\
400.00 & c1500 & -0.92 & -1.70 & -0.24 & 0.95 & 0.40\\
440.00 & c1500 & -1.26 & -2.41 & -0.18 & 0.95 & 0.28\\
480.00 & c1500 & -1.47 & -3.36 & 0.43 & 0.95 & 0.23\\
520.00 & c1500 & -1.60 & -4.86 & 1.58 & 0.95 & 0.20\\
560.00 & c1500 & -1.71 & -7.01 & 3.37 & 0.95 & 0.18\\
600.00 & c1500 & -1.88 & -10.07 & 5.98 & 0.95 & 0.15\\
240.00 & i500 & 2.31 & 0.79 & 3.93 & 0.95 & 10.10\\
280.00 & i500 & 0.55 & -0.46 & 1.52 & 0.95 & 1.72\\
320.00 & i500 & -0.54 & -1.13 & 0.08 & 0.95 & 0.58\\
360.00 & i500 & -1.08 & -1.50 & -0.61 & 0.95 & 0.34\\
400.00 & i500 & -1.22 & -1.78 & -0.65 & 0.95 & 0.30\\
440.00 & i500 & -1.08 & -2.03 & -0.19 & 0.95 & 0.34\\
480.00 & i500 & -0.81 & -2.16 & 0.59 & 0.95 & 0.44\\
520.00 & i500 & -0.55 & -2.50 & 1.42 & 0.95 & 0.58\\
560.00 & i500 & -0.42 & -3.16 & 2.28 & 0.95 & 0.65\\
600.00 & i500 & -0.58 & -4.35 & 3.10 & 0.95 & 0.56\\
240.00 & i1000 & 3.12 & 1.66 & 4.58 & 0.95 & 22.68\\
280.00 & i1000 & 1.06 & 0.15 & 1.95 & 0.95 & 2.88\\
320.00 & i1000 & -0.24 & -0.78 & 0.31 & 0.95 & 0.78\\
360.00 & i1000 & -0.92 & -1.30 & -0.52 & 0.95 & 0.40\\
400.00 & i1000 & -1.11 & -1.61 & -0.59 & 0.95 & 0.33\\
440.00 & i1000 & -0.95 & -1.80 & -0.12 & 0.95 & 0.39\\
480.00 & i1000 & -0.58 & -1.86 & 0.70 & 0.95 & 0.56\\
520.00 & i1000 & -0.14 & -1.90 & 1.77 & 0.95 & 0.87\\
560.00 & i1000 & 0.24 & -2.33 & 2.75 & 0.95 & 1.27\\
600.00 & i1000 & 0.42 & -3.17 & 3.85 & 0.95 & 1.52\\
240.00 & i1500 & 3.30 & 1.63 & 4.98 & 0.95 & 27.07\\
280.00 & i1500 & 1.08 & 0.05 & 2.14 & 0.95 & 2.94\\
320.00 & i1500 & -0.33 & -0.94 & 0.36 & 0.95 & 0.72\\
360.00 & i1500 & -1.06 & -1.52 & -0.57 & 0.95 & 0.35\\
400.00 & i1500 & -1.26 & -1.88 & -0.65 & 0.95 & 0.28\\
440.00 & i1500 & -1.06 & -1.99 & -0.09 & 0.95 & 0.35\\
480.00 & i1500 & -0.59 & -2.01 & 0.88 & 0.95 & 0.55\\
520.00 & i1500 & 0.00 & -2.05 & 2.09 & 0.95 & 1.00\\
560.00 & i1500 & 0.58 & -2.23 & 3.54 & 0.95 & 1.79\\
600.00 & i1500 & 1.01 & -3.02 & 4.86 & 0.95 & 2.75\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

\end{ThreePartTable}
\end{center}

@@

\noindent F(t) = P(RT \(\leq\) t) = 1-S(t) \hfill  (4)

The discrete-time hazard function of event occurrence gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins. While the discrete-time hazard function assesses the unique risk of event occurrence associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs afer bin t. The probability mass function cumulates the risk of event occurrence in bin t with the risks of event nonoccurrence in bins 1 to t-1. From equation 3 we find that hazard in bin t is equal to P(t)/S(t-1).

@@

which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977). This latter function is also known as the micro-level speed-accuracy tradeoff (SAT) function.

The survivor function provides a context for the hazard function, as S(t-1) = P(RT \textgreater{} t-1) = P(RT \textgreater= t) tells you on how many percent of the trials the estimate h(t) = .. is based. The probability mass function provides a context for the conditional accuracy function, as P(t) tells you on how many percent of the trials the estimate ca(t) is based.

When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition. Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t). The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet.

@@

\begin{itemize}
\tightlist
\item
  censor(df,timeout,bin\_width) : divide the time segment (0,timeout{]} in bins, identify any right-censored observations, and determine the discrete RT (time bin rank)
\item
  ptb(df) : transform the person-trial data set to the person-trial-bin data set
\item
  setup\_lt(ptb) : set up a life table for each level of 1 independent variable
\item
  setup\_lt\_2IV(ptb) : set up a life table for each combination of levels of 2 independent variables
\item
  calc\_ca(df) : estimate the conditinal accuracies when there is 1 independent variable
\item
  calc\_ca\_2IV(df) : estimate the conditional accuraies when there are 2 independent variables
\item
  join\_lt\_ca(df1,df2) : add the ca(t) estimates to the life tables (1 independent variable)
\item
  join\_lt\_ca\_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
\item
  extract\_median(df) : estimate quantiles S(t).50 (1 independent variable)
\item
  extract\_median\_2IV(df) : estimate quantiles S(t).50 (2 independent variables)
\item
  plot\_eha(df,subj,haz\_yaxis) : create plots of the discrete-time functions (1 independent variable)
\item
  plot\_eha\_2IV(df,subj,haz\_yaxis) : create plots of the discrete-time functions (2 independent variables)
\item
  plot\_eha\_agg(df,subj,haz\_yaxis) : create 1 plot for aggregated data (1 independent variable)
\end{itemize}

@@

cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}ONE + \(\alpha\)\textsubscript{2}(TIME-9) + \(\alpha\)\textsubscript{3}(TIME-9)\(^2\){]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1} + \(\beta\)\textsubscript{2}X\textsubscript{2} + \(\beta\)\textsubscript{3}X\textsubscript{2}(TIME-9){]}

@@

cloglog{[}h(t){]} = ln(-ln{[}1-h(t){]}) = {[}\(\alpha\)\textsubscript{1}D1 + \(\alpha\)\textsubscript{2}D2 + \(\alpha\)\textsubscript{3}D3 + \(\alpha\)\textsubscript{4}D4 + \(\alpha\)\textsubscript{5}D5 + \(\alpha\)\textsubscript{6}D6{]} + {[}\(\beta\)\textsubscript{1}X\textsubscript{1}{]}
